---
title: "Tugas Akhir Regspas"
author: "Angga Fathan Rofiqy"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: tango
    code_folding: hide
    toc_depth: 3
    number_sections: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    fig_caption: true
pkgdown:
  as_is: true
---

Kode di `Hide` dalam *default*, untuk menampilkan kode, klik `Code` .

```{r, warning=FALSE, message = FALSE}
#                      -=( Install & Load Package Function )=-
install_load <- function (package1, ...)  {   

   # convert arguments to vector
   packages <- c(package1, ...)

   # start loop to determine if each package is installed
   for(package in packages){

       # if package is installed locally, load
       if(package %in% rownames(installed.packages()))
          do.call('library', list(package))

       # if package is not installed locally, download, then load
       else {
          install.packages(package)
          do.call("library", list(package))
       }
   } 
}

#Path Function
path <- function(){
  gsub  ( "\\\\",  "/",  readClipboard ()  )
}
#Copy path, Panggil function di console
#Copy r path, paste ke var yang diinginkan
```

```{r , include=FALSE, echo=FALSE, warning=FALSE, message = FALSE}
#Set Working Directory
require("knitr")
opts_knit$set(root.dir = "C:/Users/Fathan/Documents/Obsidian Vault/2. Kuliah/Smt 5/4. Regresi Spasial/Tugas Akhir")
```

```{r}
#Export chart
export.chart <- "C:/Users/Fathan/Documents/Obsidian Vault/2. Kuliah/Smt 5/4. Regresi Spasial/Tugas Akhir/Chart"
```

# Pendahuluan

```{r message=FALSE, warning=FALSE}
install_load("raster","spdep","sp","readxl","openxlsx","rgdal","corrplot","DescTools","nortest","car","spatialreg","tidyverse","caret","leaps","olsrr","rio","DT","dplyr")
```

## Data Preparation

### Import Data

```{r message=FALSE, warning=FALSE}
raw.data <- import("https://raw.githubusercontent.com/Zen-Rofiqy/STA1352-RegSpas/main/Tugas%20Akhir/Data.csv")
peta <- readOGR(dsn ="Sumut", layer = "Sumatera_Utara_ADMIN_BPS")
peta <- peta[-which(peta@data$Kabupaten == "Danau Toba"), ]

datatable(raw.data, filter = 'top', 
          options = list(pageLength = 5))
```

### Data Checking

> Cek Tipe data

```{r}
str(raw.data)
```

Tidak ada masalah dalam tipe data.

> Cek Data kosong

```{r}
sum(is.na(raw.data))
raw.data[!complete.cases(raw.data),]
data <- raw.data
```

Tidak ada data kosong.

# Eksplorasi

```{r message=FALSE, warning=FALSE}
install_load('ggplot2',"cowplot","ggtext")
theme1.1 <- list(
  theme(axis.text.x = element_text(angle = 45, hjust = 1, 
                                   margin = margin(b = 10, t=20)),
        axis.text.y = element_text(vjust = 0.5, face = "bold", 
                                   margin = margin(l = 50, r = 20)),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        text = element_text(size = 30),
        plot.subtitle = element_text(hjust = 0.5),
        panel.background = element_rect(fill = 'transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(linewidth = 1, colour = "black"))
        ) 
```

## Sebaran data peubah y dan x

```{r dpi=300, fig.height = 20, fig.width = 40, fig.align = "center", message = FALSE, warning=FALSE}
y <-
ggplot(data, aes(x="", 
                     y=y)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "Y") + theme1.1 

x1 <-
ggplot(data, aes(x="", 
                     y=x1)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X1") + theme1.1 

x2 <-
ggplot(data, aes(x="", 
                     y=x2)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X2") + theme1.1 

x3 <-
ggplot(data, aes(x="", 
                     y=x3)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X3") + theme1.1 

x4 <-
ggplot(data, aes(x="", 
                     y=x4)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X4") + theme1.1 

x5 <-
ggplot(data, aes(x="", 
                     y=x5)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X5") + theme1.1 

x6 <-
ggplot(data, aes(x="", 
                     y=x6)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X6") + theme1.1 

x7 <-
ggplot(data, aes(x="", 
                     y=x7)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X7") + theme1.1 

x8 <-
ggplot(data, aes(x="", 
                     y=x8)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X8") + theme1.1 

x9 <-
ggplot(data, aes(x="", 
                     y=x9)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X9") + theme1.1 

x10 <-
ggplot(data, aes(x="", 
                     y=x10)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X10") + theme1.1 

x11 <-
ggplot(data, aes(x="", 
                     y=x11)) +
  #Violin
  geom_violin(scale="count", fill='#991B1E', alpha=0.25, 
                    trim = FALSE) +
  #Boxplot
  geom_boxplot(fill='#FBCE32', color='black',
                     outlier.size=4, outlier.color='#FBCE32', 
                     notch=T, width=0.2) +
  #Mean Marker
  stat_summary(fun = mean, geom = "point", shape = 16, size = 4, 
               color = "black") +
  theme(plot.title = element_text(hjust=0.5),legend.position = "none") +
  labs(x = "\n", 
       y='',
       title = "X11") + theme1.1 

chart <-
  plot_grid(y, x1, x2, x3, x4, x5, x6, x7, x8, x9, x10, x11,
            nrow=2)
chart
```

## Boxplot of Variables

```{r}
for (i in 5:ncol(data) ) {
  boxplot(data[,i],horizontal=TRUE, main = paste("Sebaran dari ",colnames(data)[i]))
}
```

## Pairs

```{r}
library(DescTools)
pairs(data[,5:ncol(data)], pch = 19, lower.panel = NULL)
```

## Correlation Matrix

```{r}
library(corrplot)
corrplot(cor(data[,5:ncol(data)]), method = "number")
```

## Poverty Heatmap (%)

```{r}
k=101
colfunc <- colorRampPalette(c("green", "yellow","red"))
color <- colfunc(k)
peta$y <- data$y
spplot(peta, "y", col.regions=color)
```

# OLS Finding

## Initial/Prior OLS Model

```{r}
ols <- lm(y~., data = data[,5:ncol(data) ] )
summary(ols)
```

## Best Model (Forward Selection)

```{r}
optmodfw = ols_step_forward_p(ols)
summary(optmodfw$model)
```

## Best Model (Backward Selection)

```{r}
optmodbw = ols_step_backward_p(ols)
summary(optmodbw$model)
```

## Best Model (Stepwise)

```{r}
optmodsw = ols_step_both_p(ols)
summary(optmodsw$model)
```

## Best 7-Predictor (7P) Model Taken From Best Subset

```{r}
Best_Subset <-
    regsubsets(y~.,
               data = data[,5:ncol(data) ],
               nbest = 1,      # 1 best model for each number of predictors
               nvmax = 7,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "exhaustive")
summary_best_subset <- summary(Best_Subset)

summary_best_subset$which[7,2:ncol(summary_best_subset$which)]
```

## Correlation Matrix of Best 7P Model

```{r}
chosencol = append( as.vector(TRUE), as.vector(summary_best_subset$which[7,2:ncol(summary_best_subset$which)]) )
sevendata = data[ ,5:ncol(data)][,chosencol]
corrplot(cor(sevendata),method = "number",type = 'upper')
```

## Summary of Best 7P Model

```{r}
sevenols = lm(y~., data = sevendata )
summary(sevenols)
```

## VIF Chosen OLS

```{r}
vif(sevenols)
```

## AIC Chosen OLS

```{r}
AIC(sevenols)
```

## Normal Error Test

```{r}
err.sevenols <- residuals(sevenols)
ad.test(err.sevenols)
```

# Contiguity Weights

## Queen Contiguity

```{r}
sp.peta <- SpatialPolygons(peta@polygons)
qc <- poly2nb(sp.peta, queen = T)
qc
```

```{r}
W.qc <- nb2listw(qc, style='W',zero.policy=TRUE)
qct = lm.morantest(sevenols, W.qc, alternative="greater", zero.policy = TRUE)
qct
```

## Rook Contiguity

```{r}
sp.peta <- SpatialPolygons(peta@polygons)
rc <- poly2nb(sp.peta, queen = F)
rc
```

```{r}
W.rc <- nb2listw(rc, style='W',zero.policy=TRUE)
rct = lm.morantest(sevenols, W.rc, alternative="greater", zero.policy = TRUE)
rct
```

# Distance Weights

```{r}
longlat = coordinates(peta)
peta$long <- longlat[,1]
peta$lat <- longlat[,2]
coords <- peta[c("long","lat")]
koord <- as.data.frame(coords)
djarak<-dist(longlat)
m.djarak<-as.matrix(djarak)
```

## KNN

```{r}
W.knn<-knn2nb(knearneigh(longlat,k=5,longlat=TRUE))
W.knn.s <- nb2listw(W.knn,style='W')
mt1 = lm.morantest(sevenols,W.knn.s,alternative = "greater")
mt1
```

## Radial Distance Weight

```{r}
W.dmax<-dnearneigh(longlat,0,60,longlat=TRUE)
W.dmax.s <- nb2listw(W.dmax,style='W', zero.policy = TRUE)
mt2 = lm.morantest(sevenols,W.dmax.s,alternative = "greater", zero.policy = TRUE)
mt2
```

## Exponential Distance Weight

```{r}
alpha=1
W.e<-exp((-alpha)*m.djarak)
diag(W.e)<-0
rtot<-rowSums(W.e,na.rm=TRUE)
W.e.sd<-W.e/rtot #row-normalized
W.e.s = mat2listw(W.e.sd,style='W')
mt5 = lm.morantest(sevenols,W.e.s,alternative ="greater")
mt5
```

## Exponential Distance Weight ($\alpha = 2$)

```{r}
alpha2=2
W.e2<-exp((-alpha2)*m.djarak)
diag(W.e2)<-0
rtot2<-rowSums(W.e2,na.rm=TRUE)
W.e2.sd<-W.e2/rtot2 #row-normalized
W.e2.s = mat2listw(W.e2.sd,style='W')
mt6 = lm.morantest(sevenols,W.e2.s,alternative = "greater")
mt6
```

## All Spatial Weights in Comparison

```{r}
KodeMatrikBobot <- c("W.knn.s","W.dmax.s","W.e.s","W.e2.s","W.rc","W.qc")
MatrikBobot <- c("KNN","dmax","Exp1","Exp2","Rook","Queen")
IndeksMoran <- c(mt1$estimate[1],mt2$estimate[1],mt5$estimate[1],mt6$estimate[1],rct$estimate[1],qct$estimate[1])
pv = c(mt1$p.value,mt2$p.value,mt5$p.value,mt6$p.value,rct$p.value,qct$p.value)
M = cbind.data.frame(KodeMatrikBobot,MatrikBobot,IndeksMoran,"p-value"=pv)
M
```

## Choosing The Best Weight (Automatic)

```{r}
MSelect = M[which(M$`p-value` <= 0.05),]
MSelect2 = MSelect$KodeMatrikBobot[which( MSelect$IndeksMoran == max(MSelect$IndeksMoran) )]
ww = eval(str2lang(MSelect2))
```

## Chosen Weight Model (Exponential Distance with $\alpha = 2$)

```{r}
lm.morantest(sevenols, listw=ww, alternative="two.sided", zero.policy = TRUE)
```

```{r}
lmtest::bptest(sevenols)
```

# Spatial Dependence Model

```{r}
model <- lm.LMtests(sevenols,listw=ww,zero.policy = TRUE, test=c("LMerr","RLMerr","LMlag","RLMlag","SARMA"))
summary(model)
```

## SAR (FIX)

```{r}
sar <- lagsarlm(y ~., data = sevendata, listw = ww)
summary(sar, Nagelkerke = T)
```

## SEM

```{r}
#sem<-errorsarlm(y ~., data = sevendata, listw = ww)
#summary(sem)
```

## SARMA

```{r}
#SARMA <- sacsarlm(y ~., data = sevendata, listw = ww, zero.policy = TRUE)
#summary(SARMA)
```
